---
title: "R Project - Identifying individuals most likely to click an ad"
author: "Geoffrey Chege Mwangi"
date: '2022-05-27'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

### 1.1 Defining the question
- Determine individuals that are most likely to click on an ad from the observations of Exploratory Data Analysis

### 1.2 The Context
- A Kenyan entrepreneur has created an online cryptography course and would like to advertise it on her blog.
- She is targeting audiences from various countries.
- In the past, she ran ads to advertise a related course on the same blog and collected data in the process.
- She would now like to employ my services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads.

### 1.3 Metric for success
- Clear indication of which factors influence whether an individual is likely to click on an ad i.e. Gender, location, income, daily internet usage.

### 1.4 Experimental Design Taken
- Installing packages and loading libraries needed
- Loading the data
- Data Cleaning
- Exploratory Data Analysis:
  - Univariate Analysis
  - Bivariate Analysis

### 1.5 Appropriateness of the available data
- The columns in the dataset include:
  - Daily_Time_Spent_on_Site
  - Age
  - Area_Income
  - Daily_Internet_Usage
  - Ad_Topic_Line
  - City
  - Male
  - Country
  - Timestamp
  - Clicked_on_Ad

## 2. Installing and loading Necessary Packages
```{r install, eval=FALSE, include=FALSE}
install.packages("iterators")  
install.packages("caret") 
install.packages("caretEnsemble")
install.packages("ggplot2")
install.packages("e1071")
install.packages("randomForest")
install.packages("ggcorrplot")
install.packages('ranger')
install.packages('caTools')
install.packages('rpart.plot')
install.packages('psych')
```

```{r loading, include=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
library(dplyr)
library(ggplot2)
library(data.table)
library(mclust)
library(caret)
library(corrplot)
library(clustvarsel)
library(wskm)
library(cluster)
library(reshape2)
library(DataExplorer)
library(lattice)
library(Hmisc) 
library(caret)
library(psych)
library(ggcorrplot)
```

## 3. Loading the Data


```{r ad}
ad <- read.csv("C:/Users/user/Downloads/advertising.csv")
head(ad)
```


## 4. Data Cleaning

### 4.1 Checking the attribute types

```{r class, echo=FALSE}
sapply(ad,class)
```

### 4.2 converting time variable from character to date and time (POSIXct) format
```{r datetime}
ad$Timestamp <- as.POSIXct(ad$Timestamp, "%Y-%m-%d %H:%M:%S",tz = "GMT")
```

### 4.3 Checking for duplicates

```{r duplicates}
duplicates <- ad[duplicated(ad),]
duplicates
```

There are no duplicates in the dataset

### 4.4 checking for null values

```{r nulls}
colSums(is.na(ad))
```

There are no null values in the dataset

### 4.5 checking column names

```{r names}
names(ad)
```

Replacing the periods "." with underscores "_"

```{r underscore}
names(ad) <- gsub("[.]", "_", names(ad))
```

```{r name check}
names(ad)
```
### 4.6 Outliers

I will use boxplots to check for outliers.

#### Boxplot for "Area_Income"

```{r boxplot income, echo=FALSE}
boxplot(ad$Area_Income,
main = "Area Income Boxplot",
xlab = "Area Income",
col = "blue",
border = "black",
horizontal = TRUE,
notch = TRUE
)
```

#### Boxplot for "Age"

```{r boxplot age, echo=FALSE}
boxplot(ad$Age,
main = "Age Boxplot",
xlab = "Age",
col = "orange",
border = "black",
horizontal = TRUE,
notch = TRUE
)
```

#### Boxplot for "Daily_Time_Spent_on_Site"

```{r boxplot time on site, echo=FALSE}
boxplot(ad$Daily_Time_Spent_on_Site,
main = "Time spent on site Boxplot",
xlab = "Time spent on site",
col = "brown",
border = "black",
horizontal = TRUE,
notch = TRUE
)
```

#### Boxplot for "Daily_Internet_Usage"

```{r internet usage, echo=FALSE}
boxplot(ad$Daily_Internet_Usage,
main = "Daily Internet usage Boxplot",
xlab = "Daily internet usage",
col = "red",
border = "black",
horizontal = TRUE,
notch = TRUE
)
```

## 5. Exploratory Data Analysis

### 5.1 Univariate Analysis

Summary statistics of the dataset

```{r summary}
summary(ad)
```
From the summary statistics, the following can be concluded about these columns:

#### Daily_Time_Spent_on_Site:
mean: 65

median: 68.22

#### Age:
mean: 36.01

median: 35

#### Area Income:
mean: 55,000

median: 57,012

#### Daily_Internet_Usage:
mean: 180

median: 183.1

Using describe() to get range, skewness, kurtosis and standard deviation among others:
```{r inform}
describe(ad)
```

#### Mode

A function to determine the mode:

```{r mode}
mode <- function(v){
  uniq <- unique(v)
  uniq[which.max(tabulate(match(v,uniq)))]
}
```

The most recurrent Ad Topic Line:

```{r modal topic, echo=FALSE}
topic_mode <- mode(ad$Ad_Topic_Line)
topic_mode
```

The most recurrent City:
```{r modal city, echo=FALSE}
city_mode <- mode(ad$City)
city_mode
```

The most recurrent Country:
```{r modal country, echo=FALSE}
country_mode <- mode(ad$Country)
country_mode
```

Checking the modal age using a barplot:
```{r age plot, echo=FALSE}
counts <- table(ad$Age)
barplot(counts,
  main="A bar chart showing Age",
  xlab="Age",
  ylab = "Frequency",
  col=c("darkblue"),
)
```

From the plot, the modal age is 31.


Checking the distribution in terms of gender where 1 is Male and 0 is Female:
```{r gender distribution, echo=FALSE}
gender <- (ad$Male)
gender.frequency <- table(gender)
gender.frequency
# plotting to visualize the distribution
barplot(gender.frequency,
  main="A bar chart showing Gender",
  xlab="Gender",
  ylab = "Frequency",
  col=c("darkblue","red"),
  )
```

From this, there are More women than men, making female the modal gender.

### 5.2 Bivariate Analysis

#### Scatterplots

```{r}
# scatterplot
plot((ad$Daily_Time_Spent_on_Site), (ad$Age), 
     main = "A scatterplot of Time Spent on site against age",
     xlab = 'Time spent', 
     ylab = 'Age')
```

```{r}
# scatterplot of Time on site vs income
plot((ad$Daily_Time_Spent_on_Site), (ad$Area_Income), 
     main = "A scatterplot of Time Spent on site against income",
     xlab = 'Time Spent on Site', 
     ylab = 'Income')
```

```{r}
# scatterplot of Time on site vs Internet usage
plot((ad$Daily_Time_Spent_on_Site), (ad$Daily_Internet_Usage), 
     main = "A scatterplot of Time Spent on site against Daily Internet Usage",
     xlab = 'Time Spent on Site', 
     ylab = 'Daily Internet Usage')
```

#### Heatmap

```{r heatmap}
# Heat map
# Checking the relationship between the variables
# Using Numeric variables only
numeric_tbl <- ad %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage)
# Calculate the correlations
corr <- cor(numeric_tbl, use = "complete.obs")
ggcorrplot(round(corr, 2), 
           type = "full", lab = T)
```

#### Those who clicked on ads:

Analysis of people who click on the ads:

```{r}
# Analysis of people who click on the ads
ad_click <- ad[which(ad$Clicked_on_Ad == 1),]
```

Most popular age group of people clicking on ads:

```{r}
# Most popular age group of people clicking on ads
hist((ad_click$Age), 
     main = "Histogram of Age of those who click ads",
     xlab = 'Age', 
     ylab = 'Frequency',
     col = "blue")
```

40 - 45 year olds click on the most ads

#### Plotting to visualize the gender distribution:

```{r}
gender2 <- (ad_click$Male)
gender2.frequency <- table(gender2)
gender2.frequency
# plotting to visualize the gender distribution
barplot(gender2.frequency,
  main="A bar chart showing Gender of those who clicked",
  xlab="Gender(0 = Female, 1 = Male)",
  ylab = "Frequency",
  col=c("darkblue","red"),
  )
```

Females clicked more ads than males.

#### Scatterplots of those who clicked:

```{r}
# scatterplot
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Age), 
     main = "A scatterplot of Time Spent on site and clicked ad against age",
     xlab = 'Time spent', 
     ylab = 'Age')
```

```{r}
# scatterplot of Time on site vs income
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Area_Income), 
     main = "A scatterplot of Time Spent on site and ad clicked against income",
     xlab = 'Time Spent on Site', 
     ylab = 'Income')
```

```{r}
# scatterplot of Time on site vs Internet usage
plot((ad_click$Daily_Time_Spent_on_Site), (ad_click$Daily_Internet_Usage), 
     main = "A scatterplot of Time Spent on site and ad clicked against Daily Internet Usage",
     xlab = 'Time Spent on Site', 
     ylab = 'Daily Internet Usage')
```


```{r heatmap clicked}
# Heat map
# Checking the relationship between the variables
# Using Numeric variables only
numeric_tbl <- ad_click %>%
  select_if(is.numeric) %>%
  select(Daily_Time_Spent_on_Site, Age, Area_Income,Daily_Internet_Usage)
# Calculate the correlations
corr <- cor(numeric_tbl, use = "complete.obs")
ggcorrplot(round(corr, 2), 
           type = "full", lab = T)
```

The country with the most ad clicks:

```{r}
mode(ad_click$Country)
```

The income that clicks most:

```{r}
mode(ad_click$Area_Income)
```

Ad title that garners most clicks:

```{r modal topic clicked, echo=FALSE}
topic_mode2 <- mode(ad_click$Ad_Topic_Line)
topic_mode2
```

All the data profiling statistics of those who clicked on ads will be organized into the report below:

```{r include=FALSE}
create_report(ad_click)
```

The link to the report is here: file:///C:/Users/user/Documents/Geoffrey%20Chege%20Moringa%20IP%20W12/report.html


## 6. Conclusion

From the Exploratory Data Analysis, it can be concluded that those most likely to click on ads are Women from Australia, ranging from ages 40 - 45 and with an income of 24593. The ad title that is clicked on most is "Reactive local challenge".

## 7. Recommendations

- There should be more locally targeted ads, seeing as the key word 'local' prompted more clicks.